### 2014-01-30 Images/Lineups/LineupGeneration.Rnw
RNW files for stimuli to generate a single file with all stimuli.

\documentclass{article}
\usepackage{fullpage}

\begin{document}

<<setup, out.width="5in", out.height="4in", fig.width=10, fig.height=8, fig.align="center", echo=FALSE, results='asis', cache=TRUE, message=FALSE, warning=TRUE>>=
setwd("../../")
library(nullabor)
library(ggplot2)
library(plyr)
library(Hmisc)

source("./Code/theme_lineup.R")
source("./Code/GenerateData.R")

set.seed(31415926)

# Linear trend alone - no outliers
N <- 10
seeds <- c(699181, 190785, 746705, 938857, 256246, 
           130991, 829750, 107019, 310685, 968407, 
           982098, 247471, 658815, 828038, 956567, 
           706934,  74904, 743373, 955470, 339582, 
           230370, 594429, 959280, 677468, 301141, 
           523820, 397938, 952596, 726989, 247199)
# seeds <- sample(0:1e6, N) + sample(0:1e3, N) + sample(0:100, N)
answers <- data.frame()

colors <- c("#882E72", "#D92120", "#4EB265", "#3F4EA1", 
            "#E7742F", "#1965B0", "#B1B343", "#4683C1", "#DFA53A")

pics.set <- as.character(1:(N*2))
pics.set[nchar(pics.set)==1] <- paste("0", pics.set[nchar(pics.set)==1], sep="")
names <- data.frame(filename = sample(paste("./Images/Lineups/", rep(1:3, each=N*2), "-", rep(pics.set, times=3), ".png", sep=""), 6*N), idx=1:(6*N), stringsAsFactors=FALSE)
names$title <- gsub(".png", "", gsub("./Images/Lineups/", "", names$filename, fixed=TRUE), fixed=TRUE)

plots <- as.list(rep(NA, length(names$filename)))
names(plots) <- names$filename

permute.var <- c("x", "y")
for(i in 1:N){
  j <- 3*i-2
  set.seed(seeds[i])
  dframe <- linear.trend()

  # Slope alone
  pos.x <- sample(1:20, 1)
  filename <- names$filename[j]
  pv <- sample(permute.var, 1)
  lineupdata <- lineup(null_permute(pv), dframe, pos=pos.x)
  plots[[j]] <- ggplot(data=lineupdata) + 
                  geom_point(aes(x=x, y=y), size=3) + 
                  facet_wrap(~.sample) +
                  theme_lineup() + 
                  ggtitle(names$title[j])
  answers <- rbind(answers, data.frame(name=filename, target1=pos.x, target2=NA, seed=seeds[i], idx=i, permute.var=pv, group="NoOutliers", type="Slope"))

  # Color alone
  pos.x <- sample(1:20, 1)
  pv <- "group"
  filename <- names$filename[j+1]
  lineupdata <- lineup(null_permute(pv), dframe, pos=pos.x)
  plots[[j+1]] <- ggplot(data=lineupdata) + 
                  geom_point(aes(x=x, y=y, color=factor(group)), size=3) + 
                  facet_wrap(~.sample) + 
                  scale_color_manual(values=colors[c(2,8)]) +
                  theme_lineup()+ 
                  ggtitle(names$title[j+1])
  answers <- rbind(answers, data.frame(name=filename, target1=NA, target2=pos.x, seed=seeds[i], idx=i, permute.var=pv, group="NoOutliers", type="Color"))

  # Slope vs. Color
  pos.x <- sample(1:20, 1)
  pos.y <- sample(c(1:20)[which(1:20!=pos.x)], 1)
  pv <- sample(permute.var, 1)
  filename <- names$filename[j+2]
  lineupdata <- permute.groups2(lineup(null_permute(pv), dframe, pos=pos.x), 
                                ngroups=3, pos=pos.y)

plots[[j+2]] <- ggplot(data=lineupdata) + 
                geom_point(aes(x=x, y=y, color=factor(group.k)), size=3) + 
                facet_wrap(~.sample) + 
                scale_color_manual(values=colors[c(3,5,6)]) +
                theme_lineup() + 
                ggtitle(names$title[j+2])
  answers <- rbind(answers, data.frame(name=filename, target1=pos.x, target2=pos.y, seed=seeds[i], idx=i, permute.var=pv, group="NoOutliers", type="ShapeColor"))
}

# write.csv(answers, "./Images/Lineups/NoOutliersKey.csv", row.names=FALSE)

# Linear trend with Outliers
N <- 10
set.seed(271828)
seeds <- c(488462,  89012,  68718, 875098, 790630,
           995942, 307213, 635549, 380455, 971611,
            90250, 687530, 557990, 873571, 611075,
           100338, 214593, 365595, 609200, 944618,
           578770, 710917, 117886, 458621, 307274,
           749423, 231016, 321914, 222946, 627623)
# seeds <- sample(0:1e6, N) + sample(0:1e3, N) + sample(0:100, N)
# answers <- data.frame()

colors <- c("#882E72", "#D92120", "#4EB265", "#3F4EA1", 
            "#E7742F", "#1965B0", "#B1B343", "#4683C1", "#DFA53A")
permute.var <- c("x", "y")
for(i in 1:N){
  set.seed(seeds[i])
  j <- 28+3*i
  dframe <- linear.trend(n=50, n.outliers=4, sd=4)

  # Slope alone
  pos.x <- sample(1:20, 1)
  filename <- names$filename[j]
  pv <- sample(permute.var, 1)
  lineupdata <- lineup(null_permute(pv), dframe, pos=pos.x)
  plots[[j]] <- ggplot(data=lineupdata) + 
                  geom_point(aes(x=x, y=y), size=3) + 
                  facet_wrap(~.sample) +
                  theme_lineup() + 
                  ggtitle(names$title[j])
  answers <- rbind(answers, data.frame(name=filename, target1=pos.x, target2=NA, seed=seeds[i], idx=i, permute.var=pv, group="Outliers", type="Shape"))

  # Color alone
  pos.x <- sample(1:20, 1)
  pv <- "group"
  filename <- names$filename[j+1]
  lineupdata <- lineup(null_permute(pv), dframe, pos=pos.x)
  plots[[j+1]] <- ggplot(data=lineupdata) + 
                    geom_point(aes(x=x, y=y, color=factor(group)), size=3) + 
                    facet_wrap(~.sample) + 
                    scale_color_manual(values=colors[c(3,5,6)]) +
                    theme_lineup()+ 
                    ggtitle(names$title[j+1])
  answers <- rbind(answers, data.frame(name=filename, target1=NA, target2=pos.x, seed=seeds[i], idx=i, permute.var=pv, group="Outliers", type="Color"))

  # Slope vs. Color
  pos.x <- sample(1:20, 1)
  pos.y <- sample(c(1:20)[which(1:20!=pos.x)], 1)
  pv <- sample(permute.var, 1)
  filename <- names$filename[j+2]
  lineupdata <- permute.groups2(lineup(null_permute(pv), dframe, pos=pos.x), 
                                ngroups=4, pos=pos.y)

  plots[[j+2]] <- ggplot(data=lineupdata) + 
                    geom_point(aes(x=x, y=y, color=factor(group.k)), size=3) + 
                    facet_wrap(~.sample) + 
                    scale_color_manual(values=colors[c(3,5,6,1)]) +
                    theme_lineup() +
                    ggtitle(names$title[j+2])
  answers <- rbind(answers, data.frame(name=filename, target1=pos.x, target2=pos.y, seed=seeds[i], idx=i, permute.var=pv, group="Outliers", type="ShapeColor"))
}
plots <- plots[order(names(plots), decreasing=F)]
answers$name <- as.character(answers$name)
answers <- answers[order(answers$name, decreasing=F),]

for(i in 1:length(plots)){
  print(plots[[i]])
#   cat("/clearpage")
}
@

\end{document}