### 2015-06-23 Paper/InfoVis2015Revision/ResponsesToReviewers.Rmd
Responses to reviewers

---
title: "Responses To Reviewers"
author: "Susan VanderPlas"
date: "June 23, 2015"
output: word_document
---

# Major Concerns
1. Are all the experimental design choices properly motivated? (refer to
   the first concern of R1).    
*The purpose of each of the cognitive tests is described in section 2.2. In summary, the paper-folding, card rotation, and figure classification tasks are part of the kit of factor-referenced cognitive tests (and are thus extremely well-studied); additionally, the authors chose the specific visual searching task (VST) because of its similarity of form to statistical lineups and its' compatibility with the written form of the other cognitive tests and lineup tests.*    
*As mentioned in the paper, the paper folding and card rotation tests have been used in many previous studies of statistical graphics and data displays. The figure-classification task examines inductive reasoning, which is another component of the lineup task (determining the dimension on which the lineup components differ, then applying that to answer the question successfully).*
2. Are the experimental procedure sufficiently detailed? (refer to the
   second concern of R1 and )
3. The results of statistical analysis do not seem to be conclusive, are
   the interpretations of these results properly revised? (refer to the
   third concern of R1)
4. Does the revision contain additional discussion of why PCA is a
   proper approach to interpret the data for the VIS community members who
   do not have sufficient statistics background? (refer to the concern of
   R2)
5. Does the revision has the clear description of "most different"?
   (refer to the first concern of R3)
6. Does the revision has the proper discussion of visualization design
   (refer to the second concern of R3)
7. Does the revision resolve a potential concern of confounding factors
   (refer to the conern of R3)
8. Does the revision have an explicit related works section? (refer to
   the last concern of R3)
9. Does the revision handle all the minor issues raised by reviewers?